Delivering Reliable Software in the European Scientific Arena: Evolution of Methodologies and New Trend Analysis

### 2017-12-28 paper-ieee-trans.tex
Update paper-ieee-trans.tex

The E--Infrastructure for Testing, Integration and Configuration of Software
\cite{cordis:etics, cordis:etics2} (ETICS, Jan 2006 -- Feb 2010) project aimed
at addressing the challenges in producing quality software in distributed,
collaborative projects such as EGEE and its gLite middleware. The framework
integrated different technologies and tools to provide automated configuration,
build and testing capabilities, as well as auto-generated documentation and
software metrics gathering such as Source Lines Of Code (SLOC), complexity and
number of defects/bugs \cite{etics}. The ETICS portal was the first automated
service for delivering software products in distributed environments like
the Grids.

@@

The European Middleware Initiative (EMI, May 2010 -- Apr 2013)
\cite{cordis:emi} project joined the 4 major Grid middleware providers in
Europe at the time -- {\sl gLite}, {\sl UNICORE}, {\sl ARC} and {\sl dCache} --
to maintain and evolve the middleware focusing on extending the
interoperability in the Grid and improving the reliability of the services. The
ISO/IEC 9126 \cite{iso-9126} standard was used in order to identify a set of
characteristics that needed to be present in the EMI software products and
processes to be able to meet the EMI quality requirements
\cite{emi-quality-model}.

For each software characteristic, a set of associated
metrics and Key Performance Indicators (KPIs) were identified and defined in
detail in the EMI Metrics Specification \cite{emi-metrics}. The project
leveraged the {\sl ETICS} service for the development and release management,
as well as for metric tracking, making queries on the collected data to display
them through the chart generation framework.

@@

The EGI--InSPIRE (Integrated Sustainable Pan--European Infrastructure for
Researchers in Europe, May 2010 -- May 2014) project \cite{cordis:egi-inspire}
arose as the continuation of the EGEE--III project, establishing and
maintaining a sustainable European Grid Infrastructure.

@@

EGI--InSPIRE did not
develop any software but maintained a production--ready Grid computing
middleware distribution called {\sl UMD} (Unified Middleware Distribution).
The role of {\sl UMD} was to enforce the fulfilment of a set of quality
criteria definitions \cite{egi-qc} in the software being delivered (i.e.,
{\sl EMI} and {\sl Globus}) \cite{mario}. {\sl UMD} is still being used and
deployed in the European scientific e--Infrastructures under a follow--up
project, the EGI--Engage (Engaging the EGI Community towards an Open Science
Commons) \cite{cordis:egi-engage}. This distribution is currently complemented
by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).

@@

The INDIGO--Datacloud project reflects the progress made in software
quality and reliability aspects throughout the past scientific European
experiences. Nevertheless, the project’s Quality Assurance procedures are also
highly influenced by the insights of current big worldwide collaborations of
software development. These collaborations have continuously inspired new
technologies and procedures to increase the robustness and quality of the
software being delivered. In this scenario, the DevOps culture is one of the
major outcomes that has been progressively adopted in the INDIGO--DataCloud
project.

>>>>

The E--Infrastructure for Testing, Integration and Configuration of Software
\cite{cordis:etics, cordis:etics2} (ETICS, Jan 2006 -- Feb 2010) project aimed
at addressing the challenges in producing quality software in distributed,
collaborative projects such as EGEE and its gLite middleware. The framework
integrated different technologies and tools in order to provide automated configuration,
build and testing capabilities, as well as auto-generated documentation and
software metrics gathering such as Source Lines Of Code (SLOC), complexity and
number of defects/bugs \cite{etics}. The ETICS portal was the first automated
service for delivering quality software products in distributed environments like
the Grids.

@@

The European Middleware Initiative (EMI, May 2010 -- Apr 2013)
\cite{cordis:emi} project joined the 4 major Grid middleware providers in
Europe at the time -- {\sl gLite}, {\sl UNICORE}, {\sl ARC} and {\sl dCache} --
with the goal to maintain and evolve the middleware focusing on extending their
interoperability and improving the reliability of the services. The
ISO/IEC 9126 \cite{iso-9126} standard was used in order to identify a set of
characteristics that needed to be present in the EMI software products and
processes to be able to meet the EMI quality requirements
\cite{emi-quality-model}.

For each software characteristic, a set of associated
metrics and Key Performance Indicators (KPIs) were identified and defined in
detail in the EMI Metrics Specification \cite{emi-metrics}. The project
leveraged the {\sl ETICS} service for the development, continuous integration and release management,
as well as for metric tracking, making queries on the collected data to display
them through the chart generation framework.

@@

The EGI--InSPIRE (Integrated Sustainable Pan--European Infrastructure for
Researchers in Europe, May 2010 -- May 2014) project \cite{cordis:egi-inspire}
arose as the continuation of the EGEE--III project, with the mission of establishing and
maintaining a sustainable European Grid Infrastructure.

@@

EGI--InSPIRE not only ensured the coordination of the operation of a secure, reliable 
European-wide production grid infrastructure federated from national grid initiatives 
that is integrated and interoperates with other grids worldwide but also foreseed the 
maintenance and development of operational tools, like the Operations Portal [ref], 
the EGI Helpdesk [ref] or Grid configuration repository (GOCDB) [ref] and the 
coordination of the of  the external provision of software required by EGI to 
form the production infrastructure. For the later it maintained a production--ready 
Grid computing middleware distribution called {\sl UMD} (Unified Middleware Distribution) and 
defined general and component specific quality criteria to be applied to software 
components made available through this repository.
The role of {\sl UMD} was to enforce the fulfilment of this set of quality
criteria \cite{egi-qc} in the software being delivered (i.e.,
{\sl EMI} and {\sl Globus}) \cite{mario}. {\sl UMD} is still being used and
deployed in the European scientific e--Infrastructures under a follow--up
project, the EGI--Engage (Engaging the EGI Community towards an Open Science
Commons) \cite{cordis:egi-engage}. This distribution is currently complemented
by a Cloud--specific one called {\sl CMD} (Cloud Middleware Distribution).

@@

The INDIGO--Datacloud project reflects the progress made in software
quality and reliability aspects throughout the past scientific European
experiences. Nevertheless, the project’s Quality Assurance procedures are also
highly influenced by the insights of current big worldwide collaborations of
software development. These collaborations have continuously inspired the use new
technologies and procedures to increase the robustness and quality of the
software being delivered. In this scenario, the implementation of the DevOps culture 
is one of the major outcomes, as this paradigm has been progressively adopted by
project.
